---
title: "HSMC_NullModel"
author: "Adrián Barrero"
date: "25 de enero de 2022"
output: html_document
---
```{r}
library(BayesLogit)
library(coda)
library(Hmsc)
library(parallel)
library(corrplot)
```

```{r}
setwd("E:/ORDENADOR_UAM_2019/Clasificacion_ArcMap_Arb/1.Data/LastData")

library (readxl)
Data <- read_excel("Data2.xlsx")
Grid <- read_excel("Grid.xlsx")
```
###Data structure:
```{r}
StudyDesign <- data.frame(Plot= as.factor(Data$Plot), Grid= as.factor(Data$Grid), Year= as.factor(Data$Year))

X <- data.frame(Data[,4:15])
Y <- as.matrix(Data[,16:26]) # 11 Species
xycoords <- as.matrix(Grid[,3:4]) # XY Coordinates per cell
colnames(xycoords) = c("x-coordinate","y-coordinate")
rownames(xycoords) = Grid$Grid
head(xycoords)
```
#Let's transform those variables which distribution is skewed left. This will ease to detect 
#linear effects:
```{r}
X$Shrub <- log(X$Shrub + 1)
X$BNDVI <- log(X$BNDVI + 1)
X$Trees <- log(X$Trees + 1)
X$BareGround <- log(X$BareGround + 1)
X$Scrub20 <- log(X$Scrub20 + 1)
X$Walls <- log(X$Walls + 1)
X$Rocks20 <- log(X$Rocks20 + 1)
X$Cereal <- log(X$Cereal + 1)
X$Herbaceus <- log(X$Herbaceus + 1)
X$Rocks5 <- log(X$Rocks5 + 1)
X$Scrub5 <- log(X$Scrub5 + 1)
```
#Standardise all variables::
```{r}
X$Year <- as.factor(X$Year)
X$Shrub <- as.vector(scale(X$Shrub))
X$BNDVI <- as.vector(scale(X$BNDVI))
X$Trees <- as.vector(scale(X$Trees))
X$BareGround <- as.vector(scale(X$BareGround))
X$Scrub20 <- as.vector(scale(X$Scrub20))
X$Walls <- as.vector(scale(X$Walls))
X$Rocks20 <- as.vector(scale(X$Rocks20))
X$Cereal <- as.vector(scale(X$Cereal))
X$Herbaceus <- as.vector(scale(X$Herbaceus))
X$Rocks5 <- as.vector(scale(X$Rocks5))
X$Scrub5 <- as.vector(scale(X$Scrub5))
```
#### Fitting the model
```{r}
Ypois <- Y
Yprobit <- ifelse (Ypois > 0, 1, 0) # no use to pisson distribution
#con esa estructura estamos diciendo que me convierta los valores de abundancia en 1 y 0.
```

# We describe the random effects
```{r}
# rL.Year = HmscRandomLevel(units= levels(StudyDesign$Year)) # Year
rL.plot = HmscRandomLevel(units= levels(StudyDesign$Plot)) # Plot
# rL.Grid = HmscRandomLevel(units= levels(StudyDesign$Grid)) # Grid
rL.nngp = HmscRandomLevel(sData = xycoords, sMethod = 'NNGP', nNeighbours = 10) # XY coordinates (grid)
```
# We describe the structure of the fixes effects

```{r}
## Null model
XFormula.null = ~1
```
##### we adjust Script
```{r}
m = Hmsc(Y= Yprobit, XData= X, XFormula= XFormula, studyDesign= StudyDesign, ranLevels= list("Grid"= rL.nngp, "Plot"= rL.plot), distr= "probit")

setwd("C:/Users/AB.5047919/Desktop/AA")
localDir = "."
model.directory = file.path(localDir, "Models_Null")
```

#Only first time
```{r}
nChains = 2
samples = 100
for (thin in c(10,100)){
  transient = 50*thin
  
  m = sampleMcmc(m, thin = thin, samples = samples, transient = transient, nChains = nChains, updater=list(GammaEta=FALSE))
  filename=file.path(model.directory, paste0("model_pa_chains_",as.character(nChains),"_samples_",as.character(samples),"_thin_",as.character(thin)))
  save(m, file=filename)
    }
```

##Load to see differences.

#Thin 10
```{r}

```
#Thin 100_null_Estandarized
```{r}
#Thin 100/smples250/pachains4/O2
load(file = "E:/ORDENADOR_UAM_2019/Clasificacion_ArcMap_Arb/1.Data/LastData/ModelsO2_21092021/NullModel_pa_chains_4_samples_250_thin_100")

mpost.null = convertToCodaObject(m) # Convert to CODA object
plot(mpost.null$Beta[,3])

```

```{r}
ess.beta = effectiveSize(mpost.null$Beta)
hist(ess.beta, xlab = expression("Effective sample size" ~ beta ~ ""))
ess.V = effectiveSize(mpost$V)
hist(ess.V, xlab = "Effective sample size V")
```
#Gelman diagnostic
```{r}
psrf.beta = gelman.diag(mpost.null$Beta,multivariate=FALSE)$psrf
hist(psrf.beta, xlab = expression("Potential scale reduction factor" ~ beta ~ ""))
psrf.V = gelman.diag(mpost$V,multivariate=FALSE)$psrf
hist(psrf.V, xlab = "Potential scale reduction factor V")
```
# Explanatory power
```{r}
preds = computePredictedValues(m, updater= list(GammaEta=FALSE))
MF.nngp = evaluateModelFit(hM= m, predY= preds)
MF.nngp
```
```{r}
names(Data[16:26])
```
#predictive power
```{r}
partition = createPartition(m, nfolds = 2, column = "Grid")
cvpreds.nngp = computePredictedValues(m, partition=partition,updater=list(GammaEta=FALSE))
cvMF.nngp = evaluateModelFit(hM= m, predY= cvpreds.nngp)
cvMF.nngp
```

## Pred porw without partition
```{r}
partition = createPartition(m, nfolds = 2, column = "Grid")
cvpreds.nngp = computePredictedValues(m, partition=partition)
cvMF.nngp = evaluateModelFit(hM= m, predY= cvpreds.nngp,updater=list(GammaEta=FALSE))
sum(cvMF.nngp$TjurR2)
```
## Model Parameters
```{r}
summary(mpost$Beta)
```
##plot beta para plotear las relaciones anteriores
```{r}
postbeta = getPostEstimate(m, parName="Beta")
plotBeta(m, post = postbeta)
```
## SPECIES-TO-SPECIES ASSOCIATIONS
##plot
```{r}
OmegaCor = computeAssociations(m)
supportLevel = 0.95
toPlot = ((OmegaCor[[2]]$support>supportLevel)
          + (OmegaCor[[2]]$support<(1-supportLevel))>0)*OmegaCor[[2]]$mean

corrplot(toPlot, method = "color",
         col = colorRampPalette(c("blue","white","red"))(200),
         title = paste("random effect level:", m$rLNames[2]), mar=c(0,0,1,0))
```
##### Nuevo Plot Species associations at the Plot level model Null
```{r, fig.width= 10}
par(mar=c(0,0,0,0), family= 'serif')

spNames= c("Dupont's lark", "Common skylark", "Greater short-toed lark", "Tawny pipit", "Black-eared wheatear", "Northern wheatear", "Thekla lark", "Calandra lark", "Woodlark", "Rock sparrow", "Common stonechat")
m$spNames <- spNames

OmegaCor = computeAssociations(m)
supportLevel = 0.95
toPlot = ((OmegaCor[[2]]$support>supportLevel)
          + (OmegaCor[[2]]$support<(1-supportLevel))>0)*OmegaCor[[2]]$mean

corrplot(toPlot, method = "color",
         col = colorRampPalette(c("lightblue2","white","forestgreen"))(200),
         #title = paste("        Null Model
         # Random effect level:", m$rLNames[2]), 
         mar=c(0,0,1,0),
         diag= FALSE, tl.cex= 1.5, cl.cex= 1.25, tl.col= "black", cl.ratio= 0.35, cl.length= 5
         )

```
